Latitudinal gradients in biotic niche breadth vary across ecosystem types.

### 2015-09-04 manuscript/latitudepaper.tex
spellcheckered

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by using food webs comprised of ``trophic
    species'' --aggregations of species with idendical sets of predators and prey-- rather than  
    species \emph{per se}~\cite{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original (i.e., without aggregating 
    any species) and trophic-species (i.e., after aggregating species with identical predators and prey) versions
    of the dataset; in each case using the number of species and 
    feeding links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not species with identical sets of predators and prey are included. For simplicity, 
    here we present only the results for the original (un-aggregated, original species) webs.

@@

  Importantly, while terrestrial communities are also strongly seasonal at
  high latitudes and can recieve significant allochthonous
  inputs~\cite{Nakano2001}, terrestrial consumers tend to be
  morphologically specialised for feeding on particular prey~\cite{Liem1990}.
  This means that primarily gape-limited aquatic consumers tend to be more
  general across all types of aquatic habitats~\cite{Liem1990,Shurin2006}. The
  key to this explanation of the differences between freshwater and
  marine and estuarine  ecosystems is whether the former experience more
  severe seasonal variation. Although we are not aware of any study explicitly
  comparing seasonal variation in freshwater and saltwater or brackish food
  webs in a similar location, we believe that freshwater ecosystems are indeed
  likely to  experience more severe changes because of their small size. While
  oceans and estuaries certainly vary in terms of water temperature and
  nutrients over the course of a year~\cite{Baird1989}, these changes are likely to be slower
  and milder than in freshwaters because marine and estuarine communities are buffered 
  by being open to the ocean rather than isolated
  in the midst of a terrestrial matrix . Net primary productivity in particular is
  much more variable over the course of a year in non-marine communities~\cite{Field1998},
  suggesting that niche breadths may also be more variable over the course of the year.

>>>>

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by using food webs comprised of ``trophic
    species'' --aggregations of species with identical sets of predators and prey-- rather than  
    species \emph{per se}~\cite{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original (i.e., without aggregating 
    any species) and trophic-species (i.e., after aggregating species with identical predators and prey) versions
    of the dataset; in each case using the number of species and 
    feeding links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not species with identical sets of predators and prey are included. For simplicity, 
    here we present only the results for the original (un-aggregated, original species) webs.

@@

  Importantly, while terrestrial communities are also strongly seasonal at
  high latitudes and can receive significant allochthonous
  inputs~\cite{Nakano2001}, terrestrial consumers tend to be
  morphologically specialised for feeding on particular prey~\cite{Liem1990}.
  This means that primarily gape-limited aquatic consumers tend to be more
  general across all types of aquatic habitats~\cite{Liem1990,Shurin2006}. The
  key to this explanation of the differences between freshwater and
  marine and estuarine  ecosystems is whether the former experience more
  severe seasonal variation. Although we are not aware of any study explicitly
  comparing seasonal variation in freshwater and saltwater or brackish food
  webs in a similar location, we believe that freshwater ecosystems are indeed
  likely to  experience more severe changes because of their small size. While
  oceans and estuaries certainly vary in terms of water temperature and
  nutrients over the course of a year~\cite{Baird1989}, these changes are likely to be slower
  and milder than in freshwaters because marine and estuarine communities are buffered 
  by being open to the ocean rather than isolated
  in the midst of a terrestrial matrix . Net primary productivity in particular is
  much more variable over the course of a year in non-marine communities~\cite{Field1998},
  suggesting that niche breadths may also be more variable over the course of the year.
